The Processing of American Indian/Alaska Native Children through non-Tribal

### 2019-12-17 ms.rmd
update with proability models

```{r include=F}
rm(list=ls()); gc()
library(tidyverse)
source("lifetable.r")

pop<-read_csv("./data/pop_aian_nat.csv")
afcars<-read_csv("./data/afcars_subset.csv",
                 col_types = cols(stfcid="c"))

@@

### can't use FCID because of separate imputation
```

@@

## $Pr(\textsf{investigation})$

@@

Use chid to identify probability of ever foster care for those ever investigated.

Use CHID - match to AFCARS

Get first FC placement - use ncands_first counts as denominator.

could also do matching based on ID

@@

fc_tables_pop<-bind_rows(fc_tables)
```

```{r fctables.ncands}

>>>>

```{r include=F}
rm(list=ls()); gc()
### for libra
setwd("~/Projects/ai_an_transitions")
library(tidyverse)
library(lubridate)
source("lifetable.r")

pop<-read_csv("./data/pop_aian_nat.csv")
afcars<-read_csv("./data/afcars_subset.csv",
                 col_types = cols(stfcid="c")) %>% 
  filter(age<18)

@@

### can't use FCID because of separate imputation

afcars %>% group_by(age, year) %>% summarise(n=n())
```

@@

## $P(\textsf{investigation})$

@@

## $P(substantiation)$

```{r fctables.ncands}
ncands_first<-ncands %>% 
  filter(rptvictim==1) %>% 
  group_by(chid) %>% 
  filter(rptdt==min(rptdt))

ncands_first_subst_nat<-ncands_first %>% 
  group_by(chage, year) %>% 
  summarise(var = n()) %>% 
  rename(age = chage) %>% 
  left_join(pop) %>% 
  ungroup()

subst_tables<-list()
years<-unique(ncands_first_subst_nat$year)
for(i in 1:length(years)){
  subst_tables[[i]]<-make_life_table(ncands_first_subst_nat %>% 
                  filter(year==years[i]))
}

subst_tables<-bind_rows(subst_tables)
```

## $P(subst|inv)$

of all investigated, how many are ever substantiated?

from lifetables:

Because $P(inv|subst)=1$, and we've converted population incidence into annual age-adjusted probability, the ratio of events is the conditional probability

\[P(subst|inv) = \frac{d_{subst}}{d_{inv}}\]

```{r}

make_conditionals<-function(num, denom){
  out<-num %>% 
  select(age, year, d) %>% 
  rename(d_num = d) %>% 
  left_join(denom %>% 
              select(age, year, d) %>% 
              rename(d_denom = d)) %>% 
  group_by(year) %>% 
  summarise(q = sum(d_num)/sum(d_denom))
  return(out)
}

sub_inv<-make_conditionals(subst_tables, investigation_tables)

## $P(FC)$ using AFCARS

@@

fc_tables<-bind_rows(fc_tables)
```

## $P(FC)$ using NCANDS

```{r}
### look at first fc placement
ncands_first_fc <- ncands %>% 
  filter(!(is.na(afcarsid))) %>% 
  group_by(chid) %>% 
  filter(rptdt==min(rptdt)) %>% 
  select(chid, afcarsid, chage, year) %>% 
  ungroup() %>% 
  distinct()
### this matches length(unique(afcars$stfcid))

ncands_fc_nat<-ncands_first_fc %>% 
  rename(age = chage) %>% 
  group_by(age, year) %>% 
  summarise(var = n()) %>% 
  left_join(pop) %>% 
  ungroup()

fc_tables<-bind_rows(fc_tables)
```

## $P(FC|inv)$

```{r}
fc_inv<-make_conditionals(fc_tables, investigation_tables)
fc_sub<-make_conditionals(fc_tables, subst_tables)
```

CONDITIONALS ONLY WORK FROM SAME DATA - CANT ASSUME THAT p(b|a)=1 otherwise

need to go back to original afcars - rely on imputed ncands to get uncertainty, link to afcars by id, check year for investigation incident, maybe +/-1 yr for lags in placement starts

want to use the proportions to obtain information about probability. perhaps use a model to populate uncertainty? or construct CI from data - either way should be ok

CONFIRM THAT THEIR IS ONLY ONE FCID PER CHID

```{r}

temp<-ncands %>% 
  group_by(chid) %>% 
  summarise(n_fc=n_distinct(afcarsid))
